A large leading tail of atmospheric escape in the HAT-P-67 b system

### 2023-03-22 paper/paper1/ms.tex
Negligible corrections for Gaia DR3 bias and uncertainty

\subsection{Gaia DR3}
The stellar system consists of a binary with an M dwarf companion HAT-P-67B separated on-sky by $9\farcs0$ \citep{2019MNRAS.490.5088M}.  HAT-P-67A (\emph{Gaia DR3 1358614983131339392}) has a parallax of $2.69\pm0.01$ mas in \emph{Gaia} DR3, placing it at about 372$\pm$1.4 pc, slightly farther than previously estimated by \citet{2017AJ....153..211Z} using \emph{Gaia} DR1. Its coarse \emph{Gaia}-based estimates for RV and Doppler broadening are consistent with the more precise values published in \citet{2017AJ....153..211Z}.  Its wide companion HAT-P-67B (\emph{Gaia DR3 1358614983131339904}) \citep{2019MNRAS.490.5088M} has statistically identical parallax ($2.58\pm0.05$ mas) and proper motions, confirming its interpretation as co-moving, with projected separation of about 3400 AU.  The IAU naming convention would demand HAT-P-67A\emph{b} to refer to the planet.  Hereafter, we simply drop the A designation for notational simplicity, since the wide companion will not factor into our analysis.

>>>>

\subsection{Gaia DR3}
The stellar system consists of a binary with an M dwarf companion HAT-P-67B (\emph{Gaia DR3 1358614983131339904}) separated on-sky by $9\farcs0$ \citep{2019MNRAS.490.5088M}.  HAT-P-67A (\emph{Gaia DR3 1358614983131339392}) has a parallax of $2.69\pm0.01$ mas in \emph{Gaia} DR3, placing it at about 372$\pm$1.4 pc.  Minor corrections to the parallax uncertainty \citep{2021MNRAS.506.2269E} and bias \citep{2021A&A...649A...4L} appear negligible for the system properties.  The DR3 parallax places the system slightly farther than previously estimated by \citet{2017AJ....153..211Z} using \emph{Gaia} DR1-derived parallax of $2.92\pm0.23$, which included bias and  . Its coarse \emph{Gaia}-based estimates for RV and Doppler broadening are consistent with the more precise values published in \citet{2017AJ....153..211Z}.  Its wide companion HAT-P-67B  \citep{2019MNRAS.490.5088M} has statistically identical parallax ($2.58\pm0.05$ mas) and proper motions, confirming its interpretation as co-moving, with projected separation of about 3400 AU.  The IAU naming convention would demand HAT-P-67A\emph{b} to refer to the planet.  Hereafter, we simply drop the A designation for notational simplicity, since the wide companion will not factor into our analysis.
